Bow shocks, bow waves, and dust waves. III. Diagnostics

### 2019-03-25 papers/Papers1234/bs-bw-dw-03.tex
Finish sec 6.4 and half of sec 6.3 of Paper III

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/cygnus-bows-midir-combo}
  \caption{Two bows in Cygnus}
  \label{fig:cygnus-bows}
\end{figure}

>>>>

We have proposed a novel diagnostic method, the \(\tau\)-\(eta\shell\)
diagram, for analyzing observations of stellar bow shocks around OB
stars, which allows discrimination between radiation-supported and
wind-supported bows.  Our principal results are as follows:
\begin{enumerate}[1.]
\item The UV optical depth of the bow shell can be estimated from the
  observed infrared shell luminosity \(L\IR\) and stellar luminosity
  \(L_*\) as \(\tau = 2 L\IR/L_*\).
\item With the  
\end{enumerate}

@@

### 2019-03-25 papers/Papers1234/sec-observational-diagnostics.tex
Finish sec 6.4 and half of sec 6.3 of Paper III

\subsection{Why the two mass-loss methods differ}
\label{sec:underst-diff-betw}

@@

Out of the three quantities, \(R_0\), \(\ell/R_0\), and \(U\) (which
are highly correlated between themselves) we will concentrate on the
relative chord length \(\ell/R_0\) as the most likely culprit for the
discrepancy between the two mass-loss estimates.  This is because it
contains observational information that is used in the K18 method
(step K\ref{K2} of \S~\ref{app:bow-shock-data}), but is \textit{not
  used} in the \(\tau\)--\(\eta\shell\) method.  The closest
equivalent of \(\ell\) in our method is the shell thickness,
\(h\shell\), which enters in step~P\ref{P2} of
\S~\ref{sec:energy-trapp-vers}, but crucially we use a fixed fraction
of the bow radius, \(h\shell / R_0 = 0.25\), rather than basing it on
any observations.  In principle, we could derive \(h\shell\) from
\(\ell\) if we knew the radius of curvature of the shell,
\(R_{\text{c}}\), or equivalently the planitude:
\(\Pi = R_{\text{c}} / R_0\)\citep{Tarango-Yong:2018a}.  This is
illustrated in the inset to Figure~\ref{fig:H-versus-ell}, from which
we find
\begin{equation}
  \label{eq:h-R0}
  \frac{h\shell}{R_0} = \frac{1}{8\Pi} \left(\frac{\ell}{R_0}\right)^2 \ .
\end{equation}
This is plotted in Figure~\ref{fig:H-versus-ell} for planitudes of
\(\Pi = 1.8 \pm 0.9\), which are typical of OB star bows (this will be
shown in detail in the upcoming Paper~IV).  Also shown in the figure
are the individual relative chord lengths for the K18 sources (rug
plot at top).  Although the mean value of \(\ell/R_0 \approx 1.8\)
corresponds closely to the \(h\shell/R_0 = 0.25\) that we have used,

\subsection{Shared properties of candidates for radiation support}
\label{sec:comm-prop-cand}

For five of the K18 sources (380, 407, 409, 410, 411), our method of
\S~\ref{sec:mass-loss-determ} gives only an upper limit to the wind
mass-loss rate if one assumes a factor-three uncertainty in
\(\eta\shell\), raising the possibility that they may be
radiation-supported instead of wind-supported (see Paper~I).  It is
notable that these sources are among the smallest in the K18 sample,
all with \(R_0 < \SI{0.1}{pc}\).  The two Orion Nebula bows, LP~Ori
and \thD{} also fall into this category, and they are smaller still,
with \(R_0 < \SI{0.01}{pc}\).  Six of these 7 stars also occupy a
narrow range of spectral type\footnote{%
  However, this is not particularly statistically significant.  For
  instance, out of the twenty K18 sources, 7 have spectral type O8 or
  earlier.  Therefore, taking as the null hypothesis that the 5
  sources were randomly selected from the 20, then the chance that
  they should all be O9 or later is
  \(p = (1 - \frac{7}{20})^5 = 0.116\), which fails to meet the
  conventional \(2\sigma\) significance level of \(p = 0.05\).} %
between O9 and B0, corresponding to
\(T_{\text{eff}} = \text{\SIrange{28}{31}{kK}}\) (LP~Ori is a much
cooler B2 star with \(T_{\text{eff}} = \SI{20}{kK}\),
\citealp{Petit:2008a, Alecian:2013a}).

The small bow sizes are fully consistent with the expectations from
theory developed in Paper~I.  Radiation-supported bows are predicted
to occur either in high-density environments
(\(n > \SI{100}{cm^{-3}}\)) or for stars with particularly weak winds.
In either case, the bow is expected to be smaller than \SI{0.1}{pc}
for \(v_\infty \ge \SI{20}{km.s^{-1}}\), see Figures~2a,b and~6 of
Paper~I.\@

Another common property of these sources is that they all have high
ratios of chord length to stand-off radius, \(\ell/R_0\).  For
instance, from Figure~\ref{fig:H-versus-ell}, the four sources with
the highest \(\ell/R_0\) values are all among the five
radiation-supported candidates.\footnote{%
  In this case, the statistical significance is far higher, since
  \(p = (\frac{5}{20})^4 = 0.0039\) for the null hypothesis.}  From
equation~\eqref{eq:h-R0} this means they also should have thicker than
average shells.  This is consistent with our finding from Paper~I's
\S~3.3 that the shell tends to be broad in the radiation-supported bow
wave regime.  Furthermore, if the larger than average \(H/R_0\) were
to be taken account in the calculation of \(\eta\shell\)
(eq.~\eqref{eq:eta-shell}), then it would strengthen the case for
radiation support in these objects.

@@

Using emissivity at different wavelength, such as \SI{24}{\um}.  No
good reason to fear stochastic heating.  They mention not getting 70/24 ratio right.  Probably be better with our correction.

Correlations between observed parameters and with mass loss rates.
Pair plot from my jupyter notebook.

How both the UV opacity and the IR emissivity depend on the same VSG grains.

Effects of grain evolution.  RATD destroying larger grains.  How would
this effect the abundance of smaller grains? Calculate maximum grain
size for the \(U\) distribution of the K18 sample.  Do we see any
effects that seem to vary with \(U\)?

Evidence for small grains around Arches cluster near Galactic center \citep{Hankins:2017a}

>>>>

\subsection{Why do the two mass-loss methods differ?}
\label{sec:underst-diff-betw}

@@

Out of the three quantities, \(R_0\), \(\ell/R_0\), and \(U\) (which
are highly correlated between themselves) we will concentrate on the
relative chord length \(\ell/R_0\) as the most likely culprit for the
discrepancy between the two mass-loss estimates.  This is because it
contains observational information that is used in the K18 method
(step K\ref{K2} of \S~\ref{app:bow-shock-data}), but is \textit{not
  used} in the \(\tau\)--\(\eta\shell\) method.  The closest
equivalent of \(\ell\) in our method is the shell thickness,
\(h\shell\), which enters in step~P\ref{P2} of
\S~\ref{sec:energy-trapp-vers}, but crucially we use a fixed fraction
of the bow radius, \(h\shell / R_0 = 0.25\), rather than basing it on
any observations.  In principle, we could derive \(h\shell\) from
\(\ell\) if we knew the radius of curvature of the shell,
\(R_{\text{c}}\), or equivalently the planitude:
\(\Pi = R_{\text{c}} / R_0\)\citep{Tarango-Yong:2018a}.  The idealized
geometry is illustrated in the inset to Figure~\ref{fig:H-versus-ell},
from which we find
\begin{equation}
  \label{eq:h-R0}
  \frac{h\shell}{R_0} = \frac{1}{8\Pi} \left(\frac{\ell}{R_0}\right)^2 \ .
\end{equation}
This is plotted in Figure~\ref{fig:H-versus-ell} for planitudes of
\(\Pi = 1.8 \pm 0.9\), which are typical of OB star bows (this will be
shown in detail in the upcoming Paper~IV).  Also shown in the figure
are the individual relative chord lengths for the K18 sources (rug
plot at top).  Although the mean value of \(\ell/R_0 \approx 1.8\)
corresponds closely to the \(h\shell/R_0 = 0.25\) that we have used,
the sources with \(\ell/R_0 < 1\) are predicted to have thinner shells
with \(h\shell/R_0 < 0.1\).  The four sources with the lowest values
of \(\ell/R_0\) are labeled on the figure (369, 339, 382, 344) and
these are precisely those sources that show the largest discrepancy
between the two mass loss methods
(Fig.~\ref{fig:mass-loss-comparison}).  Using a smaller value of
\(h\shell\) would increase \(\eta\shell\) (eq.~[\ref{eq:eta-shell}])
and hence \(\dot{M}\) (eq.~[\ref{eq:mdot-tau-eta}]), thereby reducing
the discrepancy between the two methods.

We choose not make this correction to the \(\tau\)--\(\eta\shell\)
method in this paper since we want to test the methods using published
data alone and without excessive fine-tuning.  However, for future
applications an empirical measurement of the shell thicknesses would
clearly help improve the reliability of the method.

\subsection{Common characteristics of radiation-supported candidates}
\label{sec:comm-prop-cand}

For five of the K18 sources (380, 407, 409, 410, 411), our method of
\S~\ref{sec:mass-loss-determ} gives only an upper limit to the wind
mass-loss rate if one assumes a factor-three uncertainty in
\(\eta\shell\), raising the possibility that they may be
radiation-supported instead of wind-supported (see Paper~I).  It is
notable that these sources are among the smallest in the K18 sample,
all with \(R_0 < \SI{0.1}{pc}\).  The two Orion Nebula bows, LP~Ori
and \thD{} also fall into this category, and they are smaller still,
with \(R_0 < \SI{0.01}{pc}\).  Six of these 7 stars also occupy a
narrow range of spectral type\footnote{%
  However, the statistically significance is not strong.  For
  instance, out of the twenty K18 sources, 7 have spectral type O8 or
  earlier.  Therefore, taking as the null hypothesis that the 5
  sources were randomly selected from the 20, then the chance that
  they should all be O9 or later is
  \(p = (1 - \frac{7}{20})^5 = 0.116\), which fails to meet the
  conventional \(2\sigma\) significance level of \(p = 0.05\),
  thereby lending only weak support for rejection of the null.} %
between O9 and B0, corresponding to
\(T_{\text{eff}} = \text{\SIrange{28}{31}{kK}}\) (LP~Ori is a much
cooler B2 star with \(T_{\text{eff}} = \SI{20}{kK}\),
\citealp{Petit:2008a, Alecian:2013a}).

The small bow sizes are fully consistent with the expectations from
the theory developed in Paper~I.  Radiation-supported bows are predicted
to occur either in high-density environments
(\(n > \SI{100}{cm^{-3}}\)) or for stars with particularly weak winds.
In either case, the bow is expected to be smaller than \SI{0.1}{pc}
for \(v_\infty \ge \SI{20}{km.s^{-1}}\), see Figures~2a,b and~6 of
Paper~I.\@

Another common property of these sources is that they all have high
ratios of chord length to stand-off radius, \(\ell/R_0\).  For
instance, from Figure~\ref{fig:H-versus-ell}, the four sources with
the highest \(\ell/R_0\) values are all among the five
radiation-supported candidates.\footnote{%
  In this case, the statistical significance is far higher, since
  \(p = (\frac{5}{20})^4 = 0.0039\) for the null hypothesis, giving
  strong support for its rejection.}  From equation~\eqref{eq:h-R0}
this means they also should have thicker than average shells.  This is
consistent with our finding from Paper~I's \S~3.3 that the shell tends
to be broad in the radiation-supported bow wave regime.  Furthermore,
if the larger than average \(H/R_0\) were to be taken account in the
calculation of \(\eta\shell\) (eq.~[\ref{eq:eta-shell}]), then it
would strengthen the case for radiation support in these objects.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/cygnus-bows-midir-combo}
  \caption{Two neighboring but highly contrasting bows in Cygnus:
    KGK2010~2 (source~342) and KGK2010~1 (source~341). Each panel
    shows the same \(3.6' \times 7.4'\) field, centered on equatorial
    coordinates
    \((\alpha, \delta) = 20^{\text{h}}34^{\text{m}}34.50^{\text{s}},
    \ang{+41;58;27.4}\), but in different combinations of three
    filters, as marked. }
  \label{fig:cygnus-bows}
\end{figure}

K18 source~342 is the only source that shows a significant discrepancy
\textit{above} the theoretical mass-loss rate predictions.  This is
bow shock candidate~2 in a survey of mid-infrared arcs in Cygnus~X
\citep{Kobulnicky:2010a}, and is referred to as KGK2010~2 in
subsequent works.  For its early~B spectral type,
\(\dot{M} \approx \SI{3e-9}{M_\odot.yr^{-1}}\) is expected, whereas
the derived values from both mass-loss methods (panels a and b of
Fig.~\ref{fig:mass-loss-vs-luminosity}) are about 10 times larger.
This fact was already noted in K18, who suggested that the grain
emissivity models may not be appropriate for this object.  We propose
instead that the shell may be predominantly neutral instead of
ionized, which might help explain other peculiar aspects of the
source.  Figure~\ref{fig:cygnus-bows} shows infrared images of the
field containing both sources~342 and 341.  The field is located in
the Cygnus~X North region, between the DR~20 and DR~21 star forming
regions, about \SI{17}{pc} north of the center of the Cygnus OB2
association \citep[e.g.,][]{Schneider:2016a}.  It can be seen that the
two bow shocks have very contrasting morphologies and spectra.
Source~341 has a semi-circular shape, a very diffuse outer boundary,
and is only prominent between wavelengths of \SI{12}{\um} and
\SI{24}{\um}.  These characteristics are typical of many optically
thin bows seen in fully ionized gas, such as the Ney--Allen nebula
around \thD{} \citetext{compare the bow shape in Fig.~3a of
  \citealp{Smith:2005a}}, source~406 \citetext{HD~92607, which is
  ERO~36 in Carina, \citealp{Sexton:2015b}}, and the central star of
RCW~120 \citetext{not included in current samples, but see
  \citealp{Mackey:2015a, Mackey:2016a}}.  Source~342, on the other
hand, has a more parabolic shape, a sharply defined arc, and emits
strongly at all wavelengths from \SI{3.6}{\um} to \SI{150}{\um}
\citetext{see also Fig.~13 of \citealp{Kobulnicky:2010a}}.  The
parabolic arc shape and the broad SED are both reminiscent of LP~Ori
(more details on the Orion Nebula bows will be provide in Paper~VI),
and another bow that bears some similarities is Carina's Sickle object
\citetext{ERO~21 \citealp{Sexton:2015b}, see also
  \citealp{Ngoumou:2013a}}.  Note also the second outer arc in this
source, which emits primarily at \SI{5.8}{\um} and \SI{7.9}{\um} and
appears similar to the double-bow structure seen in IRAS~03063+5735
\citep{Kobulnicky:2012a}.

\subsection{Evolution of grain size}
\label{sec:evolution-grain-size}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/xsec-dust-ratd-ism_10}
  \caption{Change in dust opacity due to Radiative Torque Disruption
    of large grains.  Black line shows the total absorption opacity of
    Cloudy's standard ISM dust mixture.  The red line shows the result
    of simulating the RATD process by removing the grains larger than
    \SI{0.04}{\um} and distributing their mass among smaller grain
    sizes.}
  \label{fig:ratd}
\end{figure}

The properties of dust grains in bow shocks might be significantly
different from those in the general interstellar medium, particularly
because of the strong radiation field to which they are exposed, which
has effects at both the small and large ends of the grain size
distribution.  At the small end, sub-nanometer sized particles, such
as PAHs can be destroyed by hard EUV photons
\citep{Lebouteiller:2007a, Lebouteiller:2011a}.  Larger grains, on the
other hand, are most vulnerable to Radiative Torque Disruption
\citep[RATD;][]{Hoang:2018a}, see discussion in \S~6.4 of Paper~II.\@
This is important for our diagnostics because it would tend to reduce
the average grain size, while maintaining the same total grain mass,
which would have the effect of increasing the ultraviolet opacity,
while leaving the mid-infrared opacity unchanged.\footnote{The
  mid-infrared emissivity, on the other hand, would also increase
  since smaller grains tend to have higher radiative equilibrium
  temperatures.} %
We do not know the exact size distribution of fragments that would
result from the RATD process, but we can estimate its effect by
assuming that all grains with \(a > \SI{0.05}{\um}\) are transformed
into grains with \(a = \text{\SIrange{0.025}{0.05}{\um}}\).  We
implement this crudely in the Cloudy grain model
(\S~\ref{sec:grain-temp-emiss}) by setting to zero the abundance of
graphite and silicate grains in size bins 7, 8, 9, 10, while at the
same time distributing their mass equally between size bins 5, and 6,
which increases the abundance in those bins by factors of 5.19 and
4.45, respectively.  The results are shown in Figure~\ref{fig:ratd},
which shows the wavelength-dependent absorption opacity \(\kappa\) for
both the standard ISM grain mixture and this RATD-modified mixture.
As expected, the UV opacity is increased in the RATD mixture, but only
by about 50\%, whereas the near-infrared opacity is decreased and the
visual extinction becomes much steeper, which would correspond to a
small total-to-selective extinction ratio of \(R_V < 3\).

This modest increase in opacity from RATD grain processing is well
within the systematic uncertainties that we have been assuming and so
does not invalidate the results of the previous sections.  If the
larger grains were instead to break up into many small fragments, the
effect would be much larger.  If we repeat the above exercise, but
assuming the fragment size is \(< \SI{0.01}{\um}\), then we find an
increase by a factor of five in the UV opacity, but we do not feel
that this is realistic. Studies of the break-up of fast-spinning small
asteroids \citep{Hirabayashi:2015a, Zhang:2018a} show that when the
tensile strength dominates over self-gravity, then stresses are
highest in the center of the body, leading to break up into a small
number of similarly sized pieces, as observed in asteroid P/2013~R3
\citep{Jewitt:2014a}.  In the case of dust grains, the tensile
strength and angular velocity are both far higher than in the asteroid
case, but we expect a similar behavior so that the fragment size
should be predominantly within a factor of about two below the
critical size given in equation~(28) of \citet{Hoang:2018a}, as we
assumed in Figure~\ref{fig:ratd}.

% Using emissivity at different wavelength, such as \SI{24}{\um}.  No
% good reason to fear stochastic heating.  They mention not getting 70/24 ratio right.  Probably be better with our correction.

% How both the UV opacity and the IR emissivity depend on the same VSG grains.

% Evidence for small grains around Arches cluster near Galactic center \citep{Hankins:2017a}
